---
title: "Stat Inf HW4"
## author: "Richard Decal"
## date: "November 3, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## complete the following problems. Use R for any calculations/simulations. You can use LaTeX for any equations or you may write the equations by hand, scan them in, and include them in your document.

# 1. Of the 213 countries or independent economies in the world recognized by the World Bank, 19% are classified by the World Bank as being developed. Are countries with a large land area (>500,000 km2) more likely to be developed? 23% of all countries are large, and 17% of all countries are both large and developed. Use D to denote the event of a country being developed and L to denote a country being large.
$P(L) = 0.23$

$P(L\cap D) = 0.17$

$P(D) = 0.19$

## a. What is the probability that a country is either large or developed?

$P(L) +P(D) -P(L\cap D) = 0.23 + 0.19 - 0.17 = 0.25$

## b. What is the probability that a large country is developed? Compare this to the probability that any country is developed.

$\frac{P(D\cap L)}{P(L)} = 0.17/0.23 = 0.74$

## c. What is the probability that a developed country is large? Compare this to the probability that any country is large.

$\frac{P(D\cap L)}{P(D)} = 0.17/0.19 = 0.89$

## d. In terms of countries, what would it mean to say that D and L are disjoint events? Are they disjoint?

It would mean that being developed and being large are mutually exclusive conditions. Since there are countries that are both, the sets are not disjoint.

## e. In terms of countries, what would it mean to say that D and L are independent events? Are they independent?

It would mean that the conditional probability of being developed is totally unaffected by the prior knowledge that the country is also large (and vice versa). Since there is a large overlap between the two sets, this is not true.

# 2. A certain river floods every year. Suppose that the low-water mark is set at 1 and the high water mark Y has distribution function $F_Y(y)=P(Y \leq y) =1 - \frac{1}{y^2}, 1 \leq y \leq \infty$
## a. Verify that F is a cdf.

## b. Find the pdf.
I do not know how todo this

# 3. Suppose X and Y are two random variables with E(X) = 5, Var(X) = 1, E(Y) = 11, Var(Y) = 4. 
## a. Find the covariance between X and Y if they are independent.
TODO don't know

## b. Find the variance of 2X + 3Y if they are independent.
$$Var(2x+3Y)=Var(2x)+Var(3y)\\
= 2^2Var(x) + 3^2 Var(y)\\
=4*1 + 9*4 = 40$$

## c. Find the covariance between X and Y if corr(X, Y) = 0.5. 
$corr(x,y) = \frac{cov(x,y)}{\sigma_x \sigma_y}$

# 4. R Simulation: Simulate sample data from (i) independent normal, (ii) correlated normal, and (ii) independent skewed (e.g. gamma) distributions. 
```{r}
n <- 1000
norm_samp <- rnorm(n)
gamma_samp <- rgamma(n, 2)
```



## a. For each simulation, compute 1000 samples. For each sample, find a confidence interval for the mean.
```{r}
norm_samp.mean <- mean(norm_samp)
norm_samp.sd <- sd(norm_samp)
norm_samp.mean + c(-1,1) * (qt(0.95, df=n) * norm_samp.sd / n)
```

## b. Overall, how well do the confidence intervals work when assumptions are violated?(i.e. compare the coverage from simulations (ii) and (iii) to simulation (i)).
TODO

# 5. In the Lock5Data package, there are two data sets containing information on random samples of commute times in Atlanta (CommuteAtlanta) and in St. Louis (CommuteStLouis). 

## a. Find the sample size, sample mean, and standard deviation for commute times in each city.
```{r}
library(Lock5Data)
atl.mean <- mean(CommuteAtlanta$Time)
atl.sd <- sd(CommuteAtlanta$Time)
atl.n <- length(CommuteAtlanta$Time)

stluis.mean <- mean(CommuteStLouis$Time)
stluis.sd <- sd(CommuteStLouis$Time)
stluis.n <- length(CommuteStLouis$Time)
```

## b. Create a graph of the data for each city. Are there any outliers? Based on the graph and the data itself, determine if it is appropriate to perform inference on this data?
```{r}
boxplot(CommuteAtlanta$Time, CommuteStLouis$Time, names=c("Atlanta", "St. Luis"))
cat("# Atlanta outliers: ", length(CommuteAtlanta$Time[CommuteAtlanta$Time > (atl.mean + 2*atl.sd)]))
cat("# St. Luis outliers: ", length(CommuteStLouis$Time[CommuteStLouis$Time > (stluis.mean + 2*stluis.sd)]))
```
Both cities have some outliers, but less than 4% of either set. From that, the data seems to be well behaved enough to perform inference with.


## c. Find and interpret a 95% confidence interval for the difference in commute time between the two cities. Interpret the results in context.

TODO

## d. Conduct a test to determine whether the commute time in Atlanta is longer on average than the commute time in St. Louis. State the hypotheses, test statistic, critical value, and make a formal decision at the alpha = 0.05 level of significance. Show your work for these calculations, then also show how to get the information from built-in functions in R. 

$H_0$ is that there mean(Atlanta commute) = mean(St Pete commute). $H_a$ is mean(Atlanta commute)>mean(St Pete commute). 

## e. Find the p-value for the test in part d. Interpret the p-value in context.

# 6. A statistics student was interested in prices for used Mustang cars being offered for sale on an internet site. He sampled 25 cars from the website; the data are stored in the data set “MustangPrice” in the Lock5Data package.

## a. Suppose we are interested in studying the correlation between Price and Miles for the cars. Describe how you might take one bootstrap sample from this data set. 

We would sample from the set with replacement.
```{r}
n_bootstrap <- 10000
n_cars <- length(MustangPrice$Price)

```



## b. What is the bootstrap estimate of the standard error?



## c. Find a 95% bootstrap confidence interval for the correlation. Interpret the interval in context.

## d. Suppose we would like to create a 95% bootstrap confidence interval for the median price of the mustangs. Simulate 1000 bootstrap samples and graph the bootstrap distribution for the sample median price. If appropriate, construct a 95% bootstrap CI for the median price; if not appropriate, explain why.
